Left-linear rules in adhesive categories

### 2024-02-13 icalp/fusione.tex

There is, however, a further problem. In the literature on linear
rewriting systems, a central result is the a characterisation of
switch equivalence of sequences of steps based on what are called processes or
consistent permutations~\cite{Handbook, tobias}.\todo{Handbook, graph processes, tobias}
In turn this is the key
to derive some simple but fundamental properties for linear rewriting
systems, namely

\begin{enumerate}
\item \emph{consistency of switching}: when transforming
  sequences of rewriting steps into equivalent ones by switching independent steps, the
  result not depends on the order in which switchings are applied, but
  only on the associaterd permutation;
\item \emph{globality of independence}: if two independent steps are
  moved forward or backward due to the application of switchings to a
  sequence, they remain independent.
\end{enumerate}
Properties (1) and (2) allow one to deduce a canonical form for equivalence proofs 
between sequences of rewriting steps seen as concurrent computations.

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails: the
existence of a consistent permutation is only necessary but not
sufficient, hence it cannot be used for proving (1) and (2).
%
This leads to asking whether the properties above hold.
We show that actually much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weak form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality. Finally, we prove that a canonical form
for switching sequences can also be recovered.

>>>>

There is, however, a further problem. In the literature on linear
rewriting systems, a central result is the characterisation of
switch equivalence of sequences of steps based on what are called processes or
consistent permutations~\cite{Handbook, BaldanCHKS06, tobias}.\todo{Handbook, graph processes, tobias}
In turn this is the key
to derive some simple but fundamental properties for linear rewriting
systems, namely

\begin{enumerate}
\item \emph{consistency of switching}: when transforming
  sequences of rewriting steps into equivalent ones by switching independent steps, the
  result does not depend on the order in which switchings are applied, but
  only on the associaterd permutation;
\item \emph{globality of independence}: if two independent steps are
  moved forward or backward due to the application of switchings to a
  sequence, they remain independent.
\end{enumerate}
Properties (1) and (2) allow one to deduce a canonical form for equivalence proofs 
between sequences of rewriting steps seen as concurrent computations.

Unfortunately, for left-linear rewriting systems the characterisation
of switch equivalence based on consistent permutations fails: the
existence of a consistent permutation is only necessary but not
sufficient, hence it cannot be used for proving (1) and (2).
%
This leads to asking whether the properties above hold.
We show that actually much of the theory can be retained: consistency
of switching still holds, while globality of independence holds in a
weaker form, conceptually linked to the fact that fusions introduce a
form of disjunctive causality. Finally, we prove that a canonical form
for switching sequences can also be recovered.
